Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-03 paper/paper.tex
inew graph

%\newcommand{\deltacall}[3]{\delta_{#1}({#2},{#3})}
\newcommand{\scmin} {\mbox{\sf\em in}}
\newcommand{\scmnull}{\mbox{\sf\em null?}}
\newcommand{\scmpair}{\mbox{\sf\em pair?}}
\newcommand{\scmprim}{\ensuremath{\mathsf{+}}}
\newcommand{\sh}[1]{{\colorbox{gray!20}{\framebox{$#1$}}}}
\def\myvec{\mathaccent"017E } 
\newcommand{\stk}{\mbox{S}}   
\newcommand{\ID}{\mbox{$\mathbf{ id}$}}
\newcommand{\bang}{\mbox{\sc bang}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\begin{document}
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be
%over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data
%(0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

@@

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software  Engineering}{Metrics}[complexity  measures,
  performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

By Friday, February 27 2015, 23:59 (UTC-11), submit a full paper of at
most  12 pages (6  pages for  an Experience  Report), in  standard ACM
conference format, including bibliography, figures, and appendices.

@@

While  the runtime  system  of most  functional  languages includes  a
garbage  collector  (GC)  to  efficiently  reclaim  memory,  empirical
studies on Scheme~\cite{karkare06effectiveness} and, more importantly,
on Haskell~\cite{rojemo96lag}
%and Java~\cite{shaham02estimating}
programs have shown that  GCs leave uncollected a large
number of  memory objects that are  reachable but not  live (here {\em
  live} means the  object can potentially be used by  the program at a
later stage).  This results in unnecessary memory retention.

@@

\begin{figure}[t!]
  %%%%{\color{Myblue}
  \psset{unit=1mm}
  \begin{pspicture}(80,50)(0,-50)
    %\psframe(80,50)(0,-50)
    %  \begin{center}
    \begin{tabular}{@{}c@{}}
      \begin{boxedminipage}{.42\textwidth}
        {\sf
	  \renewcommand{\arraystretch}{1}{
	    \begin{uprogram}
              \UFL\
              \UNL{0} (\DEFINE\ (\length\  \pl)
	      \UNL{1}  (\SIF~(\NULLQ \ \pl) $0$
              (\PRIM\ 1\ (\CAR\  \pl))))
              \UNL{0}
	      \UNL{0}  (\DEFINE\ (\append\  \lista\ \listb)
	      \UNL{1}  (\SIF~(\NULLQ \ \lista)
	      \listb
	      \UNL{2} \hspace*{-0.1cm}(\CONS\ (\CAR\  \lista) (\append\
              (\CDR\  \lista)
              \listb))))
              \UNL{0}
              \UNL{0} (\LET\ \px\
              $\leftarrow$\ (\CONS\ $5$
              %          \UNL{1}\hspace*{1.5cm}
              (\CONS\ (\CONS\ $6$ \NIL) \NIL) \IN
	      \UNL{1} (\LET\ \py\   $\leftarrow$\  (\CONS\ $3$ \NIL) \IN
	      \UNL{2}
              (\LET\ \pz\  $\leftarrow$\  (\append\ \px\  \py)\ \IN\
              \UNL{3} (\SIF~(\NULLQ~(\CAR~(\CDR~\pz)))~$0$~$\pi$:\
              (\length\ \pz))))))
	    \end{uprogram}
        }}
      \end{boxedminipage}
      \\
      (a) Example program. \\ \\
      \begin{minipage}{.45\textwidth}
        %\hspace*{1cm}
                {\includegraphics[width=1\textwidth]{motiv-example}}
      \end{minipage}
      \\ \renewcommand{\arraystretch}{.9}{\begin{tabular}[t]{ll}
          (b) Memory graph at $\pi$. Thick edges denote live
          links. Traversal \\{\white (b)} stops at edges
          marked $\times$ during garbage collection.
      \end{tabular}}
    \end{tabular}
  \end{pspicture}
 \vspace*{-3ex}
  \caption{Example Program and its Memory Graph}\label{fig:mot-example}
\end{figure}

>>>>

%\newcommand{\deltacall}[3]{\delta_{#1}({#2},{#3})}
\newcommand{\scmin} {\mbox{\sf\em in}}
\newcommand{\scmnull}{\mbox{\sf\em null?}}
\newcommand{\scmpair}{\mbox{\sf\em pair?}}
\newcommand{\scmprim}{\ensuremath{\mathsf{+}}}
\newcommand{\sh}[1]{{\colorbox{gray!20}{\framebox{$#1$}}}}
\def\myvec{\mathaccent"017E } 
\newcommand{\stk}{\mbox{S}}   
\newcommand{\ID}{\mbox{$\mathbf{ id}$}}
\newcommand{\bang}{\mbox{\sc bang}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\begin{document}
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be
%over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data
%(0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

@@

\category{D.3.4}{Programming Languages}{Processors}[Memory Management
  (Garbage Collection), Optimizations]
\category{F.3.2}{Logic and Meanings Of Programs}{Semantics of
  Programming Languages}[Program Analysis]

\terms{Algorithms, Languages, Theory}

\keywords{Heap Analysis, Memory Management, Garbage Collection}

@@

While  the runtime  system  of most  functional  languages includes  a
garbage  collector  (GC)  to  efficiently  reclaim  memory,  empirical
studies on Scheme~\cite{karkare06effectiveness} and, more importantly,
on Haskell~\cite{rojemo96lag}
programs have shown that  GCs leave uncollected a large
number of  memory objects that are  reachable but not  live (here {\em
  live} means the  object can potentially be used by  the program at a
later stage).  This results in unnecessary memory retention.

@@

\begin{figure}[t!]
  \psset{unit=1mm}
  \begin{pspicture}(90,48)(0,-48)
    %\psframe(90,48)(0,-48)
    \begin{tabular}{@{}c@{}}
      \psframebox{\sf
	\renewcommand{\arraystretch}{1}{
	  \begin{uprogram}
            \UFL\
            \UNL{0} (\DEFINE\ (\length\  \pl)
	    \UNL{1}  (\SIF~(\NULLQ \ \pl) $0$
            (\PRIM\ 1\ (\CAR\  \pl))))
            \UNL{0}
	    \UNL{0}  (\DEFINE\ (\append\  \lista\ \listb)
	    \UNL{1}  (\SIF~(\NULLQ \ \lista)
	    \listb
	    \UNL{2} \hspace*{-0.1cm}(\CONS\ (\CAR\  \lista) (\append\
            (\CDR\  \lista)
            \listb))))
            \UNL{0}
            \UNL{0} (\LET\ \px\
            $\leftarrow$\ (\CONS\ $5$
            (\CONS\ (\CONS\ $6$ \NIL) \NIL) \IN
	    \UNL{1} (\LET\ \py\   $\leftarrow$\  (\CONS\ $3$ \NIL) \IN
	    \UNL{2}
            (\LET\ \pz\  $\leftarrow$\  (\append\ \px\  \py)\ \IN\
            \UNL{3} (\SIF~(\NULLQ~(\CAR~(\CDR~\pz)))~$0$~$\pi$:\
            (\length\ \pz))))))
	  \end{uprogram}
      }}
      \\ 
      (a) Example program. \\  \\
%      \includegraphics[width=.45\textwidth]{motiv-example}
      \input{mem-graph} \\ 
      \renewcommand{\arraystretch}{.9}
      \begin{tabular}[t]{l}
        (b) Memory graph at $\pi$. Thick edges denote live
        links. Traversal \\{\white (b)} stops at edges
        marked $\times$ during garbage collection.
      \end{tabular}
    \end{tabular} 
  \end{pspicture}
  \vspace*{-3ex}
  \caption{Example Program and its Memory Graph}\label{fig:mot-example}
\end{figure}
